---
title       : Predicting wine quality using chemical analysis
subtitle    : 
author      : Julien Grenier
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Goal

Using 2 data sets (from Cortez et al., 2009), we want to predict the wine quality (poor, standard or great) based on physicochemical attributes of the wine.

## Usage
This application (...) can be use by wine makers (with the help of a chemist) to make sure their wine contains the right chemical attributes which will yield a great quality wine.

--- .class #id 

## Data Summary

Each data sets contains 11 predictors and quality output (provided as an average of 3 wine experts). To simplify our problem we remap the original 0-10 quality scale to a 3 levels quality scale ("poor", "standard", "great").

Red wines:
```{r, echo=FALSE, cache=TRUE}
redwines <- read.csv('../winequality-red.csv', sep=';')
redwines$quality <- factor(ifelse(redwines$quality<5,"poor",ifelse(redwines$quality<7,"standard","great")))
redwines$quality <- relevel(relevel(redwines$quality,ref = "standard" ), ref="poor")
print(summary(redwines$quality))
```

White wines:
```{r, echo=FALSE, cache=TRUE}
whitewines <- read.csv('../winequality-white.csv', sep=';')
whitewines$quality <- factor(ifelse(whitewines$quality<5,"poor",ifelse(whitewines$quality<7,"standard","great")))
whitewines$quality <- relevel(relevel(whitewines$quality,ref = "standard" ), ref="poor")
summary(whitewines$quality)
```

--- .class #id 

## Model strategy

We trained 2 random-forrest models, one for each wine colors. We splitted each datasets into 2 (70% train, 30% test) and used a repeated cross-validation (10-folds 10 times) re-sampling method and used Principal Component Analysis (PCA) in pre-processing.

```{r, echo=FALSE, cache=TRUE, warning=FALSE,message=FALSE, }
library(caret)
library(doMC)
doMC::registerDoMC(4)
set.seed(54321)
trControl <- trainControl(method='repeatedcv', number=10, repeats=10, allowParallel=TRUE)

redInTrain <- createDataPartition(redwines$quality, p=0.7, list=FALSE)
redwines.training <- redwines[redInTrain,]
redwines.testing <- redwines[-redInTrain,]
redwines.model <- train(quality ~., data=redwines.training, method='rf', preProcess=c("pca"), trControl=trControl)

whiteInTrain <- createDataPartition(whitewines$quality, p=0.7, list=FALSE)
whitewines.training <- whitewines[whiteInTrain,]
whitewines.testing <- whitewines[-whiteInTrain,]
whitewines.model <- train(quality ~., data=whitewines.training, method='rf', preProcess=c("pca"), trControl=trControl)
```

```{r, echo=FALSE, cache=TRUE}
#Saving models so we don't have to train them in server.R

save(redwines.model, file="redwines.model.rda")
save(whitewines.model, file="whitewines.model.rda")
```

## Model Analysis

### Red wines
```{r, echo=FALSE, cache=TRUE}

redwines.confusionMatrix <- confusionMatrix(predict(redwines.model, newdata=redwines.testing), redwines.testing$quality)
print(redwines.confusionMatrix$table)
```
Accuracy 95% confidence interval is [`r redwines.confusionMatrix$overall[[3]]`, `r redwines.confusionMatrix$overall[[4]]`]

--- .class #id 

## Model Analysis (continued)

### White wines
```{r, echo=FALSE, cache=TRUE}

whitewines.confusionMatrix <- confusionMatrix(predict(whitewines.model, newdata=whitewines.testing), whitewines.testing$quality)
print(whitewines.confusionMatrix$table)
```
Accuracy 95% confidence interval is [`r whitewines.confusionMatrix$overall[[3]]`, `r whitewines.confusionMatrix$overall[[4]]`]